---
title: "Initial simulation explorations"
author: "Ethan C. Brown"
date: "6/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(glue)
library(psych)
load("2021-06-07 inital_sim.Rdata")

gen_unnest = gen %>% 
  unnest(cols = "sim_cell")

samp_loadings = read.delim("loadings.txt", sep = "\t", header = FALSE) %>% 
  set_names(paste0("pop_factor", 1:3))
samp_cors = read.delim("cors.txt", sep = "\t", header = FALSE) %>% 
  set_names(paste0("factor", 1:3))
rownames(samp_cors) = paste0("factor", 1:3)

factor_cutoff = 0.2
num_items = 12
```

## Methods
### Population

Population loadings:
```{r loadings}
samp_loadings %>% kable
```
Factor correlations:

```{r samp_cors}
samp_cors %>% kable
```
Possible factor score variability, as measured by item-level reliability: 0.7, 0.8, or 0.9

Ratio of sample size to # items: 25 or 50

Error SD: 1

Number of item categories: 4

Thresholds: -2.26, 0, 2.26

(Thresholds chosen so that all conditions are "unimodal" such that center two categories are at least 10% more than tail categories in all item-level reliability conditions.)

### Simulation parameters
Number of simulations per condition: 100

### Estimation
Number of factors determined by parallel analysis using polychoric correlation matrix.

Estimator is WLSMV, geomin oblique rotation.

Estimated factors matched to population factors by max congruence coefficient.

Items determined to load on factor if loading > 0.2.

## Results
### Convergence
Proportion of results that converged

### Number of factors

Number of factors detected was always correct.

```{r nfactor}
rep_nfactor = gen_unnest %>% 
  group_by(rep, r, n_multiplier) %>% 
  filter(str_starts(variable, "factor")) %>% 
  summarize(nfactor = n_distinct(variable), .groups = "drop")

rep_nfactor %>% group_by(r, n_multiplier) %>% 
  summarize(prop_correct = mean(nfactor == 3)) %>% 
  kable
```

### Did items always load on the same factors as the population?

Average proportion of pattern of loadings.

```{r gen_sort_factors}
factor_pop_sort = samp_loadings %>% as_tibble %>%
  mutate(variable = sprintf("x_%02.0f", 1:12)) %>% 
  pivot_longer(-variable, names_to = "pop_factor", values_to = "pop_loading") %>% 
  mutate(pop_loads = abs(pop_loading) > factor_cutoff) %>% 
  select(-pop_loading)

samp_pop_match = gen_unnest %>% 
  filter(!str_starts(variable, "factor")) %>% 
  pivot_longer(starts_with("factor"), names_to = "samp_factor", values_to = "samp_loading") %>% 
  mutate(samp_loads = abs(samp_loading) > factor_cutoff) %>% 
  select(-samp_loading) %>% 
  left_join(factor_pop_sort, by = "variable") %>% 
  mutate(samp_pop_match = samp_loads == pop_loads) 

match_counts = samp_pop_match %>% 
  group_by(rep, r, n_multiplier, samp_factor, pop_factor, samp_pop_match) %>% 
  summarize(match_count = n(), .groups = "drop")
  
  
max_match = match_counts %>% 
  group_by(rep, r, n_multiplier, samp_factor) %>% 
  filter(match_count == max(match_count) & samp_pop_match) %>% 
  mutate(match_prop = match_count / num_items)

match_summary = max_match %>% 
  group_by(r, n_multiplier, pop_factor) %>% 
  summarize(mean_prop = mean(match_prop), .groups = "drop") %>% 
  pivot_wider(names_from = pop_factor, values_from = mean_prop)
```
### Congruence coefficients

```{r congruence}
gen_loadings = gen_unnest %>% 
  filter(!str_starts(variable, "factor"))


samp_loadings

gen_congruence = gen_loadings %>% 
  select(-variable) %>% 
  nest_by(rep, r, n_multiplier) %>% 
  ## does taking absolute value make sense?
  mutate(congruence = list(factor.congruence(abs(as.matrix(data)),
                                             abs(as.matrix(samp_loadings))) %>% 
                             as_tibble(rownames = "samp_factor"))) %>% 
  select(-data) %>% 
  unnest(cols = congruence)
  
max_congruence = gen_congruence %>% 
  pivot_longer(starts_with("pop"),
               names_to = "pop_factor",
               values_to = "congruence") %>% 
  group_by(rep, r, n_multiplier, pop_factor) %>% 
  summarize(max_congruence_factor = samp_factor[which.max(congruence)],
    max_congruence = max(congruence), .groups = "drop")


avg_congruence = max_congruence %>% 
  group_by(r, n_multiplier, pop_factor) %>% 
  summarize(mean_congruence = mean(max_congruence), .groups = "drop") %>% 
  pivot_wider(names_from = pop_factor, values_from = mean_congruence)
  
avg_congruence %>% kable
  
  
```



