---
title: "Initial simulation explorations"
author: "Ethan C. Brown"
date: "6/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(glue)
library(psych)
load("2021-06-08 second_sim_results.Rdata")

options(knitr.kable.NA = "")

# gen_unnest = gen %>% 
#   
#   unnest(cols = "sim_cell")

samp_loadings = read.delim("loadings_2.txt", sep = "\t", header = FALSE) %>% 
  set_names(paste0("pop_factor", 1:3))
samp_cors = read.delim("cors_2.txt", sep = "\t", header = FALSE) %>% 
  set_names(paste0("factor", 1:3))
rownames(samp_cors) = paste0("factor", 1:3)

factor_cutoff = 0.2
num_items = nrow(samp_loadings)
```

## Methods
### Population

Population loadings:
```{r loadings}
samp_loadings %>% kable
```
Factor correlations:

```{r samp_cors}
samp_cors %>% kable
```
Possible factor score variability, as measured by item-level reliability: 

```{r rs}
gen %>% 
  pull(r) %>% unique %>% sort
```

Ratio of sample size to # items: 

```{r n_multiplier}
(unique_n_multiplier = gen %>% pull(n_multiplier) %>% unique %>% sort)
```

Possible numbers of items:
```{r num_items}
num_items
```

Possible sample sizes
```{r sample_sizes}
num_items * unique_n_multiplier
```

Error SD: 1

Number of item categories: 4

Thresholds: -2.26, 0, 2.26

(Thresholds chosen so that all conditions are "unimodal" such that center two categories are at least 10% more than tail categories in all item-level reliability conditions.)

### Simulation parameters
Number of simulations per condition:
```{r}
gen %>% pull(rep) %>% max
```


### Estimation
Number of factors determined by parallel analysis using polychoric correlation matrix.

Estimator is WLSMV, geomin oblique rotation.

Estimated factors matched to population factors by max congruence coefficient.

Items determined to load on factor if loading > 0.2.

## Results
### Convergence

Frequency of errors encountered:

```{r gen_is_err}
gen_is_err = gen %>% 
  rowwise() %>% 
  mutate(class = paste(class(sim_cell), collapse = " "),
         is_error = str_starts(class, "simpleError"),
         error_text = ifelse(is_error, as.character(sim_cell), "No error"),)

gen_is_err$error_text %>% table

```
Probably should use `start = FALSE` in `semTools::efaUnrotate` and skip rotation when parallel analysis only finds one factor. For now, will treat the errors about rotation as a single-factor extraction and the `start = FALSE` errors as true convergence errors.  Need to look more into the "missing value" error.

```{r convergence}
gen_is_err %>% 
  mutate(converged = !is_error | str_detect(error_text, "GPArotation")) %>% 
  group_by(r, n_multiplier) %>% 
  summarize(convergence = mean(converged), .groups = "drop") %>% 
  arrange(n_multiplier, r) %>% 
  kable
```


### Number of factors
Proportion of each number of factors extracted:
```{r nfactor}
rep_nfactor = gen_is_err %>% 
  mutate(nfactor = ifelse(is_error, 
                          ifelse(str_detect(error_text, "GPArotation"), 1, NA),
                          sum(str_starts(sim_cell$variable, "factor"))))

rep_nfactor %>%
  filter(!is.na(nfactor)) %>% 
  count(r, n_multiplier, nfactor) %>% 
  group_by(r, n_multiplier) %>% 
  mutate(proportion = n/sum(n)) %>% 
  select(-n) %>% 
  pivot_wider(names_from = nfactor,
              values_from = proportion) %>% 
  arrange(n_multiplier, r) %>% 
  kable

```

